D3064R0: Bringing the Laws of Physics, Computer-System Constraints, and Compilers to Bear on the OOTA Problem

### 2024-01-28 ootavacuous.tex
Apply hardware-architect feedback to load/store ordering appendix
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

However, if that load suffers a cache miss, address translation will have
completed long before the load returns its value, meaning that a later
store might well commit before the load completes.
In fact, that later store might commit before the store that supplies the
value returned by that load!
Which explains why some older hardware systems really do reorder prior
loads and later stores.

@@

One possible explanation is ECC errors.

If correctable ECC errors were fixed up in hardware, then only uncorrectable
ECC errors would be directly visible to software.
If the only possible reaction to an uncorrectable ECC error was to terminate
the program suffering that ECC error, there would be no problem.

However, some code has ways of handling even uncorrectable errors.
Operating-system kernels encountering an ECC error might note that the
affected data was being used by only one thread, and might react by
killing that thread.
Some user applications might be note that the affected data affected only
a particular computation, and might react by restarting that computation.
Both the kernel and the user application would need to continue execution,
and might thus desire an exact exception.

Systems that offload processing of correctable ECC errors to software
also require exact exceptions.
After all, the software is going to need to be able to correct the
error and then fix up state as if the load had returned the fixed-up
value.

>>>>

However, if that load suffers a cache miss, address translation will have
completed long before the load returns its value, meaning that a later
store might well commit before the load completes.
In fact, that later store might commit before the store that supplies the
value returned by that load!
Which explains why some hardware systems really do reorder prior loads
and later stores.

@@

One explanation is ECC errors.

If correctable ECC errors were fixed up in hardware, then only uncorrectable
ECC errors would be directly visible to software.
If the only possible reaction to an uncorrectable ECC error was to
terminate the program suffering that ECC error, it would be safe to
allow subsequent stores to commit as soon as all prior loads' address
translations had completed successfully.

However, some kernels and applications have ways of handling even
uncorrectable errors.
Operating-system kernels encountering an ECC error might note that the
affected data was being used by only one user process, and might react by
killing that user process.
Some user applications might be note that the affected data affected only
a particular computation, and might react by restarting that computation.
In both cases, it is only reasonable to react to an uncorrectable ECC
error in a load from a read-only mapped file by re-reading that data
from that file, then restarting that load instruction.
In all cases, the kernel and the user application would need to continue
execution, and would thus require an exact exception.

Systems that offload processing of correctable ECC errors to software
also require exact exceptions.
After all, the software is going to need to be able to correct the error
and then fix up the state to make it appear that the load had returned
the fixed-up value.
